---
title: "midcab_demographics"
author: "Salil V Deo"
date: "4/18/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report for initial data cleaning for the **midcab** paper.

```{r get libraries}

library(easypackages)
libraries(c("tidyverse","pastecs","skimr",
           "readxl","readr","survival","relsurv","lubridate","gmodels"))

proc_desc <- function(x){
  # require library(pastecs)
  result <- round(stat.desc(x),2)
  hist(x, col = "skyblue")
  result
}

proc_tab <- function(x){
  # require libary(gmodels)
  CrossTable(x)
}


# na converter

c.na <- function(df,x,y){
  df$x[df$x == y]<- NA
}
```



```{r}

# get data
# this the sheet 1 for the file 

df <- read_excel("C:\\github_rcode\\midcab\\data\\base.xlsx", sheet = 1)

dim(df)

glimpse(df)

# create data into sections according to the information
# Pat_ID is the patient indicator
# create table demo with baseline demographics first and then clean that


demo <- c("lfdNr", "THG_Referenz", "Aufnahme_Nr", "Pat_ID", "date_of_surgery", 
  "dob", "age_at_surgery", "gender", "admission", "height_cm", 
  "weight_kg", "BMI", "BSA_Mosteller", "pre_med_ass", "pre_med_clopidogrel", 
  "pre_med_Gp2b_3A_antagonist", "pre_med_nitrate_iv", "pre_med_inotropika", 
  "diabetes", "art_hypertension", "pulm_hypertension>60mmHg", "smoker", 
  "hyperlipemia", "COPD", "peripheral_vascular_disease", "pre_creatinine_mg_dl", 
  "pre_dialysis", "preGFR_CockcroftGault", "pre_LVEF", "pre_LVEF_function", 
  "pre_heartrhythm", "PM_wearer", "ICD_wearer", "pre_NYHA", "pre_CCS", 
  "pre_ASA", "priority", "elective", "urgent", "emergency", "critical_preoperative_state", 
  "pre_reanimation", "pre_ventilation", "coronary_disease_detail", 
  "left_main_disease", "pre_lyse", "prior_PCI", "last_PCI_date", 
  "pre_MI", "pre_MI<48hours", "pre_MI_2_to_<21days", "pre_MI_21_to_<91days", 
  "pre_MI_>90days", "pre_CVA", "pre_neurological_dysfunction", 
  "pre_cardiac_shock", "pre_cardiac_shock<48hours", "pre_iabp", 
  "pre_ecmo", "typ_a_dissection", "diss_iatrogen", "active_endocarditis", 
  "postinfarct_septal_rupture", "prior_cardiac_surgery", "last_cardiac_surgery_date", 
  "additive_euroSCORE", "logistic_euroscore", "CABG", "OPCAB", 
  "MIDCAB", "TECAB")

demog <- df[,c(demo)] # demog now has all the demographic information for all the patients

# check if the rows are unique

duplicates <- demog[duplicated(demog$Pat_ID),] # duplicated numbers 

# duplicate patient id

duplicates %>% select(Pat_ID) 
```

There is 1 duplicate patient. This is the demographics table; all that is the same. So I have kept 1 row and removed the other in table *demog2*.


```{r}
demog2 <- demog %>% 
  group_by(Pat_ID) %>%
  filter(!duplicated(Pat_ID)| n() == 1)

```

To confirm that all unique in table *demog2*

```{r}
duplicates <- demog2[duplicated(demog2$Pat_ID), ]

dim(duplicates)

```

Now to see the data and work on cleaning.
*demog2* now has 71 col. That is going to be the master df now.

```{r}

# get the first 10 col

d <- demog2[, 1:10]

glimpse(d)

```

```{r}
# age 

proc_desc(d$age_at_surgery)

# age ok ; no missing 

# convert dates to date in R 

d$date_of_surgery <- as_date(d$date_of_surgery)

d$dob <- as_date(d$dob)

# gender

str(d$gender)

table(d$gender, useNA = 'ifany') # no missing gender 1 = male, 2= female

# conv to female

d$female <- d$gender - 1

d$female <- factor(d$female, levels = c(0,1), labels = c("no","yes"))

proc_tab(d$female)

# height_cm

proc_desc(d$height_cm) # missing = -9

# convert that to missing data

d$height_cm[d$height_cm == -9]<- NA

proc_desc(d$height_cm) 

```


now go for the next 10 variables 

```{r}
d2 <- demog2 %>% select(Pat_ID,11:20)

glimpse(d2)
```


```{r}

# weight

d2$weight_kg[d2$weight_kg == -9]<- NA

proc_desc(d2$weight_kg)

# BMI

d2$BMI[d2$BMI == -9]<- NA

proc_desc(d2$BMI)

# make BMI meaningful so if BMI > 45, then converted to 45

d2$BMI[d2$BMI > 45]<- 45

proc_desc(d2$BMI)

# antiplatelet agent before surgery
# either aspirin, clopidogrel or other G2P3 antagonist

d2$antiplt <- with(d2, ifelse((pre_med_ass == 1|pre_med_clopidogrel == 1| pre_med_Gp2b_3A_antagonist == 1), 1, 0))

proc_tab(d2$antiplt)

# diabetes

proc_tab(d2$diabetes)

d2$diabetes[d2$diabetes == -9]<- NA

proc_tab(d2$diabetes)

```


now to tackle col 21:30 of the data

```{r}
d3 <- demog2 %>% select(Pat_ID, 21:30)

glimpse(d3)
```


