---
title: "midcab_datacleaning"
author: "Salil V Deo"
date: "4/18/2020"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Report for initial data cleaning for the **midcab** paper.

```{r get libraries}

library(easypackages)
libraries(c("tidyverse","pastecs","skimr",
           "readxl","readr","survival","relsurv","lubridate","gmodels","haven", "tidylog"))



proc_desc <- function(x){
  # require library(pastecs)
  result <- round(stat.desc(x),2)
  hist(x, col = "skyblue")
  result
}


proc_tab <- function(x){
  # require libary(gmodels)
  CrossTable(x, format = "SPSS")
  
}


# na converter


input_na <- function(x, y){
  ifelse(x == y, NA, x)
}
```



```{r}

# get data
# this the sheet 1 for the file 


df <- read_excel("C:\\github_rcode\\midcab\\data\\new_table.xlsx", sheet = 1)

dim(df)

glimpse(df)

#- to see if any duplicates exist

duplicated_id <- df %>% 
  group_by(Pat_ID) %>%
  filter(duplicated(Pat_ID)| n() != 1) %>%
  select(Pat_ID)

duplicated_id
# the duplicated rows are because they have different last contact dates
# am going to remove duplicates and keep the last instance of follow-up time of duplicates here.
# arrange according to id and then last contact date

df2 <- df %>% arrange(Pat_ID, last_contact_date)

# now going to remove duplicates in patient id as they are unique

df3 <- df2[!rev(duplicated(rev(df2$Pat_ID))), ]

#- df3 now has the single row which has the more recent contact date

#- so df3 is now the master

```


table df3 is the master now.



```{r}

# get the first 10 col

d <- df3

# glimpse(d)

```

got more data for tecab and cpb midcab. 
need to remove tecab patients
also recode for cpb_midcab

```{r}

# remove tecab first

tecab <- read_excel("C:/github_rcode/midcab/data/tecab.xlsx", sheet = 1)

glimpse(tecab)

tecab.l <- tecab$tecab

d$tecab_add <- with(d, ifelse(Pat_ID %in% tecab.l, 1, 0))

proc_tab(d$tecab_add)

d$tecab_actual <- with(d, if_else((TECAB ==1 |tecab_add == 1), 1, 0))

proc_tab(d$tecab_actual)

# now to clear the cpb code 

cpb <- read_excel("C:/github_rcode/midcab/data/cpb.xlsx", sheet = 1)

cpb.l <- cpb$cpb

# check if cpb and tecab are different

tecab$diff <- with(tecab, if_else(tecab %in% cpb.l, 1, 0))

table(tecab$diff) 

# clear cpb now

d$cpb_add <- with(d, if_else(Pat_ID %in% cpb.l, 1, 0))

table(d$cpb_add)

# now to add these cpb patients to the earlier coded conversion patients
# also to remove tecab patients 

d$cpb_actual <- with(d, if_else((conversion_to_onpump == 1|cpb_add == 1), 1, 0))

proc_tab(d$cpb_actual)

# cpb_actual now contains all patients who had cpb conversion

# remove tecab - tecab actual

d <- d %>% filter(tecab_actual == 0)

proc_tab(d$cpb_actual)

```








```{r}
# age 

proc_desc(d$age_at_surgery)

# age ok ; no missing 

# convert dates to date in R 

d$date_of_surgery <- as_date(d$date_of_surgery)

d$dob <- as_date(d$dob)

# gender

str(d$gender)

table(d$gender, useNA = 'ifany') # no missing gender 1 = male, 2= female

# conv to female

d$female <- d$gender - 1

d$female <- factor(d$female, levels = c(0,1), labels = c("no","yes"))

proc_tab(d$female)

# height_cm

proc_desc(d$height_cm) # missing = -9

# convert that to missing data

d$height_cm[d$height_cm == -9]<- NA

proc_desc(d$height_cm) 

```


now go for the next 10 variables 

```{r}
d2 <- d

# glimpse(d2)
```


```{r}

# weight

d2$weight_kg[d2$weight_kg == -9]<- NA

proc_desc(d2$weight_kg)

# BMI

d2$BMI[d2$BMI == -9]<- NA

proc_desc(d2$BMI)

# make BMI meaningful so if BMI > 45, then converted to 45

d2$BMI[d2$BMI > 45]<- 45

proc_desc(d2$BMI)

# antiplatelet agent before surgery
# either aspirin, clopidogrel or other G2P3 antagonist

d2$antiplt <- with(d2, ifelse((pre_med_ass == 1|pre_med_clopidogrel == 1| pre_med_Gp2b_3A_antagonist == 1), 1, 0))

proc_tab(d2$antiplt)

# diabetes

proc_tab(d2$diabetes)

d2$diabetes[d2$diabetes == -9]<- NA

proc_tab(d2$diabetes)

```


now to tackle col 21:30 of the data

```{r}
d3 <- d2

# glimpse(d3)
```

make changes to the important variables

```{r}
# copd

proc_tab(d3$COPD) # no missing in COPD

proc_tab(d3$hyperlipemia) # convert to na

d3$hyperlipemia[d3$hyperlipemia == -9]<- NA

proc_tab(d3$hyperlipemia)

proc_tab(d3$peripheral_vascular_disease)

d3$pad <- input_na(d3$peripheral_vascular_disease, -9)

proc_tab(d3$pad) # pad


proc_desc(d3$pre_creatinine_mg_dl)

d3$creat_pre <- input_na(d3$pre_creatinine_mg_dl, -9)

proc_desc(d3$creat_pre) # significant # missing

proc_tab(d3$pre_dialysis)

d3$pre_dialysis <- input_na(d3$pre_dialysis, -9)

proc_tab(d3$pre_dialysis) # 14 missing 


# lvef function

proc_tab(d3$pre_LVEF_function)

d3$pre_LVEF_function <- input_na(d3$pre_LVEF_function, -9)

proc_tab(d3$pre_LVEF_function)

#- not going to use lvef as not that important

```

move forward with variables

```{r}

d4 <- d3

# glimpse(d4)

```

```{r}

# pre_heartrhythm

proc_tab(d4$pre_heartrhythm)

d4$pre_heartrhythm <- input_na(d4$pre_heartrhythm, 9)

d4$pre_heartrhythm <- input_na(d4$pre_heartrhythm, -9)

proc_tab(d4$pre_heartrhythm)

d4$pre_af <- with(d4, ifelse(pre_heartrhythm == 2, 1, 0))

proc_tab(d4$pre_af)

# aicd/pacemaker 

proc_tab(d4$PM_wearer)

proc_tab(d4$ICD_wearer)

d4$icd_pm <- with(d4, ifelse((PM_wearer == 1|ICD_wearer == 1), 1, 0))

proc_tab(d4$icd_pm) # has either icd or pacemaker

# NYHA class 

proc_tab(d4$pre_NYHA) # lot of data missing here, cannot use

# same story with CCS 

proc_tab(d4$priority)

d4$preop_status <- factor(d4$priority, levels = c(1,2,3),
                          labels = c("elective", "urgent", "emergent"))

proc_tab(d4$preop_status)

```
move further down 

```{r}

d5 <- d4

glimpse(d5)
```


```{r}

# prior PCI

proc_tab(d5$prior_PCI)

d5$prior_PCI <- input_na(d5$prior_PCI, -9)

proc_tab(d5$prior_PCI)

#  prior MI

proc_tab(d5$pre_MI)

d5$pre_MI <- input_na(d5$pre_MI, -9)

proc_tab(d5$pre_MI)

# prior stroke

proc_tab(d5$pre_CVA)

# prior IABP

table(d5$pre_iabp) # no one had preop IABP

# prior cardiac surgery

proc_tab(d5$prior_cardiac_surgery)

```

```{r}

glimpse(d5)

```


- now this table d5 contains all the cleaned data for all the demographic variables.


----------------------------------------------------------------
```{r}

# data contains only midcab patients
# write this data as demographics cleaned

write_csv(d5, 'C:/github_rcode/midcab/data/d_demo.csv')
```

d_demo is now master.

```{r}
# get the d_demo data into the script

d <- read_csv("c:/github_rcode/midcab/data/d_demo.csv")

dim(d)

# glimpse(d)


#- am going to work with the op details now


```


```{r}

# these are the important op details that we need to look at
# coversion to CPB is provided in the other more recent table 
# cpb is the new excel sheet with contains patient id for all patients operated with CPB

opdetails <- c( "conversion_to_onpump", "onpump_beating_heart", 
"incompl_revascularisation",  
"initial_thoracotomy", "conversion_to_full_sternotomy", "length_of_surgery", 
"bypass_time", "cross_clamp_time", "cardiopleg_type", "cardiopleg_tech", 
"cardiopleg_ml", "min_temperature")

```

have already data on cpb conversion, get other important data

```{r}

# conversion_to_full_sternotomy

proc_tab(d$conversion_to_full_sternotomy)

# cpb time

proc_desc(d$bypass_time) # if cpb time = 0 then na

d$bypass_time <- with(d, input_na(bypass_time, 0))

proc_desc(d$bypass_time)

# cross-clamp time

proc_desc(d$cross_clamp_time) 


```

All patients with midcab got lima - lad, so dont have consider that.

```{r}

# these are the postoperative data col.

c("post_low_cardiac_output", 
"post_iabp", "post_ecmo", "post_MI", "post_reanimation", "post_cardiac_arrhythmia", 
"post_redo_heart", "post_rethorax_bleeding", "post_symptomatic_transitory_psychotic_syndrome", 
"post_CVA", "post_CVA>72hours", "post_CVA_rankin>0", "post_DSWI", 
"post_sepsis", "post_GI_complications", "post_maxCreatinine_mg_dl", 
"post_maxGFR_CockcroftGault", "post_dialysis", "EK", "TK", "FFP", 
"ventilation_time_until_first_extubation", "ventilation_time_total", 
"post_respiratory_failure", "re_intubation", "tracheotomie", 
"post_heartrhythm", "post_new_PM", "post_new_ICD", "dayspost", 
"ICU_1stLOS", "ICU_LOS", "n_ICU", "Re_ICU", "IMCU_1stLOS", "IMCU_LOS", 
"n_IMCU", "Re_IMCU", "discharge", "died", "died_inhouse", "died_fu", 
"survival_days=", "last_contact_date", "S_Datum", "Todesursache", 
"ReOP1_Date", "ReOP1_Text", "ReOP2_Date", "Schlaganfall", "Herzinfarkt", 
"Revask.LAD bereich 1=AcB, 2=PTCA,Stent 0=nichts", "AP CCS", 
"Revaskul.nichtLAD", "Date nicht LAD", "ReOP5_Text", "lad_area", 
"rcx_area", "rca_area", "Graft1", "Graft2", "Graft3", "Graft4", 
"Graft5", "LIMA", "RIMA", "BIMA", "Radialis", "Vene", "sonstGraft", 
"Anz_Jump_Seq_Anastomosen", "Anz_aortaleAnastomosen", "Anz_YT_Anastomosen", 
"LIMA_proximal", "RIMA_proximal", "Radialis_proximal", "anzahlArteriellerGrafts", 
"lima_lad", "lima_freegraft", "rima_freegraft", "lad_graft")
```

routine important postop events are:
post_iabp, post_MI, post_rethorax_bleeding, post_CVA, post_maxCreatinine_mg_dl, post_dialysis, dayspost, discharge, died_inhouse


```{r}

# from the list above, go 1 at at time...

proc_tab(d$post_iabp)

proc_tab(d$post_CVA)

proc_tab(d$post_MI)

d$post_MI <- with(d, input_na(post_MI, -9))

proc_tab(d$post_MI)

proc_desc(d$post_maxCreatinine_mg_dl)

d$post_maxCreatinine_mg_dl <- with(d, input_na(post_maxCreatinine_mg_dl, -9))

proc_desc(d$post_maxCreatinine_mg_dl) # creatinine too many missing,would not use

proc_tab(d$post_dialysis)

d$post_dialysis <- with(d, input_na(post_dialysis, -9))

proc_tab(d$post_dialysis)

proc_desc(d$dayspost) # LOS

d$discharge_date <- as_date(d$discharge) # discharge date

proc_tab(d$died_inhouse) # died in hospital 

proc_tab(d$died) # total died overall

proc_tab(d$died_fu) # died during follow-up

```

have a look at the long-term follow-up data 

```{r}

long_term <- c("discharge", "died", "died_inhouse", "died_fu", 
"survival_days=", "last_contact_date", "S_Datum", "Todesursache", 
"ReOP1_Date", "ReOP1_Text", "ReOP2_Date", "Schlaganfall", "Herzinfarkt", 
"Revask.LAD bereich 1=AcB, 2=PTCA,Stent 0=nichts", "AP CCS", 
"Revaskul.nichtLAD", "Date nicht LAD")

```

get the data organised for the last follow-up time
first focus on long-term survival
then look at the data for the long-term adverse events



```{r}
# correct with all changes made in original excel sheet

time <- read_excel("C:/github_rcode/midcab/data/time_fup.xlsx", sheet = 1)

glimpse(time)

d2 <- left_join(d, time, by = "Pat_ID") # new master with follow-up time added

```

d2 is the master.


```{r}

# need to create the era of surgery

str(d2$date_of_surgery)

d2$date_of_surgery <- as_date(d2$date_of_surgery)

glimpse(d2$date_of_surgery)


d3 <- d2 %>%
  dplyr::mutate(surg_year = lubridate::year(date_of_surgery), 
                surg_month = lubridate::month(date_of_surgery), 
                surg_day = lubridate::day(date_of_surgery))

d3 <- d3 %>% select(-surg_month, -surg_day)

d3 %>% count(surg_year) # see how many cases per year

# now seperate into 2 periods
# period 1 = 1997 - 2003
# period 2 = 2004 - 2010
# period 3 = 2011 - 2017

d3$period <- with(d3, if_else(surg_year %in% c(1997:2003), 1,
                                               if_else(surg_year %in% c(2004:2010), 2, 3)))

d3 %>% count(period)


write_csv(d3, "C:/github_rcode/midcab/data/d3_master_04182020.csv")

```

now the data d3 contains all the variables and data needed for actual analysis
*d3* is master
*d3* contains only *midcab* without any tecab and the latest followup time as *time* col
